The pollination trade-off

### 2018-02-27 publication/manuscript.Rmd
Add prefix for references to the supplementary info

---
title: "Pollen competition"
author: "Fernando Cagua, Hugo Marrero, Jason Tylianakis, Daniel Stouffer"
bibliography: references.bib
csl: ecology-letters.csl
output: 
  bookdown::pdf_document2: 
    keep_tex: yes
    number_sections: false
toc: false
header-includes:
  - \usepackage{booktabs}
  - \usepackage{setspace}
  - \usepackage{xr}
  - \externaldocument{supp_info}
---

@@

First, we explored wether animal pollination resulted in a quantitative increase in the pollination service.
To do that, we compared the amount of conspecific pollen on stigmas from flowers open to animal pollination and that from flowers that were bagged, where only autogamus self-pollination ocurred.
The difference between these two counts is not directly equivalent to the pollen deposited by animals because self-pollination is an adaptive trait [@Kalisz2003]. 
Nevertheless, this difference still provides valuable indication of what the animal contribution to reproductive output.
We performed this comparisons using a set of non parametric Wilcoxon tests both at the community and the the species level (grouping species across communities).

We then examined the extent to which the pollination service was influenced by the number of pollination partners while accounting for other metrics that describe the plant's pollination niche and community composition. 
To do so, we fitted two sets of generalised linear models (GLM) with deposited pollen density (pollen count per stigma) as the response variable. 
The first model focused on the conspecific pollen deposited on stigmas while the second model focused on the heterospecific pollen. 
In this approach, the coefficients of both models as well as their relationship, provide an indication of the quantity and the quality of animal pollination as well as the trade-off between these two.

In both models we expect animal pollination to be

@@

Over all study sites, we found that `r round(prop_sig_global*100)`% of species observed a significant increasse (at the $\alpha = 0.05$ level) in pollen density when comparing open and bagged flowers (Table \@ref(tab:table-species-gain-global)). 
When looking at the differences within sites, we found that only `r round(prop_consistency*100)`% of the species that were present in more than one site had the same response across community (Table \@ref(tab:table-species-gain-site); Figure \@ref(fig:fig-conspecific-per-site) and \@ref(fig:fig-conspecific-per-plant)). 
Despite these differences, the community had a comparatively small effect on wether animal pollination increased the pollen density or not (Table \@ref(tab:table-variation-across-sites)). 
When comparing the performance of the models constructed at the community or study-wide level we found that the gain in conspecific pollen was better predicted by the community level models, while the gain in heterospecific pollen was better predicted by the study-wide models (Table \@ref(S-tab:table-global-vs-community-models).

>>>>

---
title: "The pollination trade-off"
author: "Fernando Cagua, Hugo Marrero, Jason Tylianakis, Daniel Stouffer"
bibliography: references.bib
csl: ecology-letters.csl
output: 
  bookdown::pdf_document2: 
    keep_tex: yes
    number_sections: false
toc: false
header-includes:
  - \usepackage{booktabs}
  - \usepackage{setspace}
  - \usepackage{xr}
  - \externaldocument[S-]{supp_info}
---

@@

First, we explored wether animal pollination resulted in a quantitative increase in the pollination service.
To do that, we compared the amount of conspecific pollen on stigmas from flowers open to animal pollination and that from flowers that were bagged, where only autogamus self-pollination ocurred.
The difference between these two counts is not directly equivalent to the pollen deposited by animals because self-pollination is an adaptive trait [@Kalisz2003]. 
Nevertheless, this difference still provides valuable indication of what the animal contribution to reproductive output.
We performed this comparisons using a set of non parametric Mann-Whitney tests both at the community and the the species level (grouping species across communities).

Our main aim was to examine the extent to which the pollination service was influenced by the number of pollination partners while accounting for other metrics that describe the plant's pollination niche and community composition. 
To do so, we fitted two sets of linear mixed models (R package `nlme 3.1-131`, @nlme2017] focusing on conspecific and heterospecific pollen respectively. 
In both models, our response variable was the difference between pollen density (pollen counts per stigma) in open and bagged flowers. 
The coefficients obtained in both sets as well as their relationship, provide an indication of the quantity and the quality of animal pollination as well as the trade-off between these two.

Each set was comprised by models of one twhousand bootstrap resampled replicates of the depositoin data. 
We used bootstrap replicates because, as open and bagged flowers were not paired, this allowed us to model the *gain* in pollen density rather than density itself. 
This strategy, not only simplified the model specification and intepretation, but also allowed us to easily evaluate the uncertainity of the model effects. 
In each replicate we randomly sampled a number of open flowers within each community-species combination and an equal number of closed flowers. 
We then log-transformed densities to improve normality of the data and minimise the impact of outliers (Figure \@ref(S-fig:fig-deposition-logtransformed-or-not) and \@ref(S-fig:fig-deposition-gain-logtransformed-or-not)). 
Finally, we calculated the difference between the pollen density among treatmemts. 
This difference was the response variable in all our models. 
As predictors, we included the species/degree (number of animal partners), the plants' relative abundance and the temporal niche overlap [others variables will be added]. 
To calculate the plant's relative abbundance we simply aggregated the flower counts in a community over the study period. 
As with the deposition data, we found that log-transforming the abundance counts drastically improved normality (Figure \@ref(S-fig:fig-abundance-logtransformed-or-not)).
We then the log-transformed abundance counts for each month between November 2010 and March 2011 to calculate the temporal niche overlap between plants. 
Specifically, we used the Pianka's niche overlap index [@Pianka1973; `spaa 0.2.2` R package @Zhang2016] which calculates an overlap index for each pair of species in the community. 
To calculate an overlap metric per species, we simply averaged the pairwise indices for a species. 
To faciliate comparison across model estimates, we scaled all independent variables to have a mean of zero and a standard deviation of one. 
In both set of models we evaluated a set of random effects which included plant species, community, wether the site was in an agricultural or restored fragment, and the locality. 
We selected the best random structure by comparing the Akaike Information Criterion (AIC) of evaluated models fitted using restricted maximum likelyhood.

Aditionally, we repeated the procedure but instead of calculating the independent variables for each plant species at the community level, we did so across sites. 
This served two purposes. 
First, and most importantly, it allowed us to examine wether indeed community level information is necessary to explain the observed patterms or whether less granular data is sufficient. 
Second, it had the added benefit that it allowed us to include some species for which not all data necessary to calculate the independent variables was available at the community level. 
We assessed

@@

Over all study sites, we found that `r round(prop_sig_global*100)`% of species observed a significant increasse (at the $\alpha = 0.05$ level) in pollen density when comparing open and bagged flowers (Table \@ref(S-tab:table-species-gain-global)). 
When looking at the differences within sites, we found that only `r round(prop_consistency*100)`% of the species that were present in more than one site had the same response across community (Table \@ref(S-tab:table-species-gain-site); Figure \@ref(S-fig:fig-conspecific-per-site) and \@ref(S-fig:fig-conspecific-per-plant)). 
Despite these differences, the community had a comparatively small effect on wether animal pollination increased the pollen density or not (Table \@ref(S-tab:table-variation-across-sites)).

We found that the random structure that was best supported the data was consistently the one that included plant species grouped in communities as random effect (Table \@ref(S-tab:table-aic-random-models)).
When comparing the performance of the models constructed at the community or study-wide level we found that the gain in conspecific pollen was better predicted by the community level models, while the gain in heterospecific pollen was better predicted by the study-wide models (Table \@ref(S-tab:table-global-vs-community-models).
